---
title: "Species Distribution Models for Prioritization in Denmark"
author: "Derek Corcoran"
date: "`r format(Sys.time(), '%d/%m, %Y')`"
output:
  bookdown::github_document2:
    fig_caption: true
    toc: true
bibliography: biblio.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = F,
  error = F,
  warning = F
)
```

<!-- badges: start -->
<!-- badges: end -->

The goal of SpeciesDistributionModelsDanmark is to explore and generate Species distribution models for de prioritization of Denmark based on its biological diversity and landuse.

# Taxonomic cleaning

First we will read a file with all the taxa found in [arter.dk](https://arter.dk/landing-page), on the 21st of September of 2022.

for this we will need the follwing packages [@Wickham2022Readxl; @Chamberlain2020Taxize; @Chamberlain2017RGBIF]:

<details style=\"margin-bottom:10px;\">
<summary>Load packages</summary>
```{r LoadPackagesTaxonCleaning, eval = T}
library(readxl)
library(taxize)
library(rgbif)
library(janitor)
```
</details>

We first read the file with all the presences:

```{r}
Taxa <- readxl::read_xlsx("2022-09-21.xlsx") |> 
  janitor::clean_names() |> 
  dplyr::select(videnskabeligt_navn)
```

This file has `r nrow(Taxa)` entries, however it only has `length(unique(Taxa$videnskabeligt_navn))` unique entries in the attribute *videnskabeligt_navn*

## Cleaning using Taxize

First we generate a new data frame considering only the unique *videnskabeligt_navn*:

```{r}
NewTaxa <- data.frame(Taxa = sort(unique(Taxa$videnskabeligt_navn)), score = NA, matched_name2 = NA) |> 
  tibble::rowid_to_column(var = "TaxaID")
```

and then we clean it using taxize first

```{r, eval = F}
for(i in 1:nrow(NewTaxa)){
  try({
    Temp <- taxize::gnr_resolve(NewTaxa$Taxa[i],
                                         data_source_ids = "11", canonical = TRUE, best_match_only = T) |> 
      dplyr::select(score, matched_name2)
    NewTaxa[i,3:4] <- Temp
      if((i %% 100) == 0){
      print(paste(i, "of", nrow(Data1), "Ready!", Sys.time()))
      readr::write_csv(NewTaxa, "Cleaned_Taxa_Taxize.csv")
    }
    gc()
  })
  
}



```


# Presence download

# Presence cleaning



