---
title: "Species Distribution Models for Prioritization in Denmark"
author: "Derek Corcoran"
date: "`r format(Sys.time(), '%d/%m, %Y')`"
output:
  bookdown::github_document2:
    fig_caption: true
    toc: true
bibliography: biblio.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = F,
  error = F,
  warning = F
)
```

<!-- badges: start -->
<!-- badges: end -->

The goal of SpeciesDistributionModelsDanmark is to explore and generate Species distribution models for de prioritization of Denmark based on its biological diversity and landuse.

# Taxonomic cleaning

First we will read a file with all the taxa found in [arter.dk](https://arter.dk/landing-page), on the 21st of September of 2022.

for this we will need the follwing packages [@Wickham2022Readxl; @Chamberlain2020Taxize; @Chamberlain2017RGBIF]:

<details style=\"margin-bottom:10px;\">
<summary>Load packages</summary>
```{r LoadPackagesTaxonCleaning, eval = T}
library(readxl)
library(taxize)
library(rgbif)
library(janitor)
```
</details>

We first read the file with all the presences:

```{r}
Taxa <- readxl::read_xlsx("2022-09-21.xlsx") |> 
  janitor::clean_names()
```

This file has `r nrow(Taxa)` entries, however it only has `length(unique(Taxa$videnskabeligt_navn))` unique entries in the attribute *videnskabeligt_navn*

# Presence download

# Presence cleaning

